---
title: "Assignment1"
output: html_document
---
## load the data
```{r, echo=TRUE}
act=read.csv("/Users/yant/tingfen/coursera/datascience/5_reproducible/assignment1/activity.csv")
dim(act)
head(act, n=20)
```
## What is mean total number of steps taken per day
### Calculate the total number of steps taken per day
```{r, echo=TRUE}
library(dplyr)
group1=group_by(act,date)
tab=summarise(group1,sum(steps,na.rm=T))
colnames(tab)=c("data", "steps")
tab[1:2,]
tb=summarise(group1,mean(steps,na.rm=T))
tb[1:3,]
##histgram
hist(tab$steps, xlab="steps", main="Histogram of total steps per day")
mn=mean(as.numeric(as.character(tab$steps)),na.rm=T)
medn=median(as.numeric(as.character(tab$steps)), na.rm=T)
```
The mean and median of the total number of steps taken per day are `r mn` and `r medn`.

##What is the average daily activity pattern?

```{r, echo=TRUE}
group2=group_by(act,interval)
tab2=summarise(group2,mean(steps,na.rm=T))
head(tab2)
colnames(tab2)=c("interval","average_steps")
plot(tab2, type="l")
tab3=tab2[order(tab2$average_steps, decreasing = T),]
head(tab3)

```
The 830-835 5-minutes interval has the maxima averge steps, 206.1698. 
##Imputing missing values
```{r, echo=TRUE}
##check no NA
ok=complete.cases(act)
##get the number of NA
num=sum(!ok)

```
The total number of missing values in the dataset is `r num`.
###Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
day=unique(as.character(act$date))
tbl=data.frame()
for (i in 1:length(day)) {
act_d=subset(act,date %in% day[i] )
act_d[1:3,]
##using mean of the day to fill in the missing data
mn=mean(as.numeric(act_d$steps), na.rm=T)
mns=ifelse(is.na(mn),0, mn)
act_d$newSteps=ifelse(is.na(act_d$steps),mns, act_d$steps)
act_d[1:2,]
tbl=rbind(tbl,act_d)

}
tbl=tbl[,c(4,2,3)]
tail(tbl)

```
###Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
group3=group_by(tbl,date)
tab4=summarise(group3,sum(newSteps,na.rm=T))
tab4[1:3,]
colnames(tab4)=c("data", "newSteps")
hist(tab4$newSteps, xlab="total number of steps", main="Histgram of total number of steps taken on each day" )
mn1=mean(tab4$newSteps)
medn=median(tab4$newSteps)
```
The mean and median total number of steps taken per day after filling the missing data "r mn1" and "r medn".

##Activity patterns between weekdays and weekends
```{r, echo=TRUE}
tbl$date=as.Date(tbl$date)
#create a vector of weekdays
wkdays=c("Monday","Tuesday", "Wednesday", "Thursday", "Friday")
tbl$wDay=factor((weekdays(tbl$date) %in% wkdays), levels=c(FALSE, TRUE), labels = c("weekend", "weekday"))

#subset data by weekday
wkd=subset(tbl, wDay %in% "weekday")
group_wkd=group_by(wkd, interval)
tab_wkd=summarize(group_wkd, mean(newSteps))
colnames(tab_wkd)=c("interval", "newSteps")
tab_wkd$Weekday=rep("weekday",288)
head(tab_wkd)
#subset data by weekend
wkend=subset(tbl, wDay %in% "weekend")
group_wkend=group_by(wkend,interval)
tab_wkend=summarize(group_wkend,mean(newSteps))
colnames(tab_wkend)=c("interval","newSteps")
tab_wkend$Weekday=rep("weekend", 288)
head(tab_wkend)
mrg=rbind(tab_wkd, tab_wkend)
head(mrg)
##make a plot
library(lattice)
xyplot(newSteps ~ interval | Weekday, data=mrg, layout=c(1,2),type="l")
```